
# R for Reproducibility


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

## Learning Objectives

Before we begin to jump into additional tools that R can help us with to be work more efficiently and in a more reproducible manner, it is helpful to first discuss why we should consider R in the first place. After completing this section you will be able to:

- Explain why R can be especially helpful for transparent and reproducibility data analyses
- Recognize that R has a very active and supportive community and locate access points to that community
- Compare R to other similar statistical and data analysis tools and programming languages
- Describe the unique benefits of R

## Why R

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Why R?"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g21a84b32106_0_9")
```

R is a programming language and environment for statistical computing and for creating plots and graphics that was developed in 1991 by Ross Ihaka and Robert Gentleman at the University of Auckland, New Zealand. However, countless contributors have made R what it is today. 

There are some especially useful aspects about R that make it a great option for creating reproducible data analyses. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Why is R useful for Reproducibility? 1.It is free and open source, 2. The community, 3. It is designed for data wrangling and stats"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.gcf1264c749_0_135")
```

## It is free and open source

The first is that R is free and [open source](https://opensource.com/resources/what-open-source). 

The term open source means that the code is publicly available.
Thus all of the code involved in creating R is actually publicly available! This enables users to check what code is used in a particular package (a set of code that allows you to do various things) so that they can modify or build upon the code if the would like to. 

In fact, many users create their own R packages to share their code with others.  There are places such as the Comprehensive R Archive Network ([CRAN](https://cran.r-project.org/)) and elsewhere that allow users to publish their own packages for others to use. 

<div class = "dictionary">
- R programming language and enivornment for statisical computing and for creating plots and other graphics
- open source -  Code is publicly available.
- open development - Anyone can contribute to developing more R code.
- R package - A set of code that can be shared between users.


</div>

Why are these aspects good for reproducibility? 

- Since R is free, it is accessible to anyone. Therefore, anyone could run your code if you shared it with them, without them needing to buy software.
- Since R is open source, if you use packages from others, people can determine what underlying code your code used (if you tell them what version you used - more on that later!)

## The community

R has a very rich and active community!

This makes it easier to reach out to others for help, find support, find tutorials, and more. 

This means that overall your code has a better chance of being more accessible than if it were written in a language that is not open source or that has limited support. 

You can also find support to make sure your code does what you want it to, as well as support to make your code as reproducible as possible.

There are several R community groups that are especially helpful:

- [R Ladies](https://www.rladies.org/) - a support group that is not just for ladies, but is open to anyone who wants to improve their R skills! There are local chapters in many large cities that often have in-person meetings. 
- There are lots of useful resources, such as the [R for Data Science book](https://r4ds.had.co.nz/) (written by two developers at Posit (formally called RStudio) which develops lots of core R packages), resources and online courses from the [Johns Hopkins Data Science Lab](https://jhudatascience.org/courses.html) including [Open Case Studies](https://www.opencasestudies.org/), resources and workshops from [Data Carpentry](https://datacarpentry.org/),  [Dataquest](https://www.dataquest.io/v2/), and more!

See this [link](https://jhudatascience.org/intro_to_r/resources.html) for more R resources.

## Designed for Wrangling and Stats

R is a statistical programming language



