# R Markdown Reports

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Learning objectives are to be able to: Explain why R Markdown files are useful for creating data analysis reports. Recognize how to test  pieces of code within R Markdown files. Recognize how to create a full R Markdown file report in html and other formats.Explain that R Markdowns generate reports from code in a way that is a bit closer to someone else trying your code."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g20eecbcf66d_84_0")
```

## Reports support reproducibility

Using R Markdown files helps you to create supports that can more transparently show what you did for your analysis and it can help you to test that your code works as expected. Scripts allow you to save code, but they do not allow you to have the following additional benefits. 

The following are reasons why R Markdown files help reproducibility:

- They allow you to show and share your code and the output of your code (this can be done in several ways depending on what you want)
- They allow you to test if your code works outside of what is active in your environment
- They allow you to test sections and all previous sections of your code out to troubleshoot 
- They help you understand what might be wrong with your code in smaller sections of code if you have an issue

## Getting Started with R Markdown

OK, so now we know how to make a new R Markdown file from the previous section.

<details><summary> Click here for a review on how to create R Markdown files in RStudio. </summary>

To open a new R Markdown file by go to `File` --> `New File` -->`R Markdown` in the upper menu of RStudio.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Creating an R Markdown file in R Studio by selecting File, New File, R Markdown."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g20ed7630a13_1_127")
```

Creating an R Markdown file starts with a pop-up and you can simply click the OK button.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Creating an R Markdown file starts with a pop-up and you can simply click the OK button."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g20ed7630a13_1_169")
```

The new pane will open on the upper left.


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The new pane will open on the upper left."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g20ed7630a13_1_176")
```

This pane is where we can write code to save in our R Markdown report.

Thus the lower left pane is where we can test out code, but the top pane is where we can write code that we wish to save. 

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The new pane will open on the upper left and is for writing code we want to save, while the bottom pane is for writing code we want to test."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g20ed7630a13_1_186")
```

The top pane where we save code is called the editor. The lower pane where we test code is called the console.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The editor in the top left, also called source, is where we write code we want to save. The lower left is the console where we do quick tests of our code."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g20ed7630a13_1_198")
```

</details>

Now let's discuss how to start writing code in such a file.

## Writing code in R Markdown files

At the top of an R Markdown file you will see some special code that is called [YAML](https://en.wikipedia.org/wiki/YAML) code. It is commonly used to configure programming projects. It does the same for our R Markdown reports. A major difference between R and YAML is that spacing really matters for YAML.

What do we mean by configure? Configuration in programming generally refers to setting things up. 

<div class = "dictionary">

- **YAML** - A language that helps set things up and shows up by defualt when you open an R Markdown file. It is written between the two `---`.
- **Configuration** - A setup for a programming project.

</div>

Here we can see what the top of an R Markdown file looks like after we first open one.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The YAML code at the top of R Markdown files sets up how the file will look and be rendered.  Here we see that title specifies a title, author lists an author name, date will list a date, and output will specify how we want the file report to be rendered.  The three dashed lines indicate the start and end of the YAML section."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g1fa1583c827_0_16")
```




You can modify the `"Untitled"` text after `title:` to specify the title of your report. 





