
# Setting up your project

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g21a84b32106_0_13")
```



## Understand why project organization is key to reproducible analyses

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g2fea8805c08_0_1121")
```

## Understand general principles of project organization

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g2fea8805c08_0_421")
```

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g2fea8805c08_0_426")
```

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g2fea8805c08_0_442")
```

https://github.com/jhudsl/reproducible-r-example

## Navigate file paths

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MNHf8JpolaEP_vQ_kB-1xRBF9wo3haCArRu117hBoHA/edit#slide=id.g2fea8805c08_0_1337")
```

## Handy R Tools

### here package

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">The only two things that make <a href="https://twitter.com/JennyBryan?ref_src=twsrc%5Etfw">@JennyBryan</a> 😤😠🤯. Instead use projects + here::here() <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/GwxnHePL4n">pic.twitter.com/GwxnHePL4n</a></p>&mdash; Hadley Wickham (@hadleywickham) <a href="https://twitter.com/hadleywickham/status/940021008764846080?ref_src=twsrc%5Etfw">December 11, 2017</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

https://here.r-lib.org/

https://www.tidyverse.org/blog/2017/12/workflow-vs-script/

### RProj files

https://bookdown.org/ndphillips/YaRrr/projects-in-rstudio.html

### ProjectTemplate

http://projecttemplate.net/

### R Markdown files

https://rmarkdown.rstudio.com/articles_intro.html

### Quarto files

https://quarto.org/docs/get-started/hello/rstudio.html
